---
title: "Wrapper functions for data.table"
author: "kleu046"
date: "2021-08-22"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{function_details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>"
)
```

```{r setup}
library(wr.data.table)
```

# Introduction

While ***data.table*** offers blazing fast functions, the syntax may be hard to remember if not used frequently particularly to those who does not use data.table frequently.

This package aims to provide *wrapper functions* to carry out common tasks using functions in the data.table library without having to remember the data.table syntax.

This vignette outlines detail usages for these wrapper function

## dt_keeprownames()

data.table(dt) removes row names.

This function converts a data.frame into a data.table and use the row names as the first column.

To use row names as the first column, user must use data.table(df, keep.rownames = True)

where df is a data.frame object

dt_keeprownames(df) will keep the row names as the first column of the data.table


```{r keeprownames}
dt <- dt_keeprownames(data1)
dt
```

## Selecting (sel_cols, desel_cols) and renaming columns (rn_cols)

User can rename columns of the data.table using rn_cols().  Note that rn_cols() function renames the original data.table therefore assigning the output back to dt is not necessary.  The function does provide the data.table as output so that user can use this function in piping.

```{r rename_rn_column}
dt |> rn_cols(rn = `vehicle model`)
dt
```

User can select colummns using sel_cols()

```{r select_columns}
subset <- dt |> sel_cols(`vehicle model`, mpg:hp, gear, vs)
subset
```

Instead of selecting columns, user can also choose to deselect certain columns.

```{r deselect_columns}
(dt |> desel_cols(gear, vs, am, drat, carb, qsec))
```

## Filtering data using filter_rows()

User can use expressions and grepl() in filter_rows() function to filter and subset the data in rows.

```{r}
dt |> filter_rows(mpg >= 20)

dt |> filter_rows(grepl("^M", `vehicle model`))
```

## Summarizing data using set_group() and condense

set_group() function adds a "group" attribute to the input data.table.

When the condense() function is called, the group attribute is used as the "by" arugment in the data.table to summarize the data by group

The example below calculates the mean and standard deviation of the data by groups of "gear" and "vs".  Note that .N notation used in data.table also works here

```{r}
subset |> set_group(`vehicle model`, gear, vs) |> attributes()

subset |> set_group(`vehicle model`, gear, vs) |> condense(mu_mpg=round(mean(mpg),2), sd_mpg=round(sd(mpg),2), count=.N) |> rm_group()
```

## Creating new columns

User can def new columns using the def_cols() function

For example, user can create an index column followed by using the arrange_cols() function to place the index column at the front of the data.table

```{r}
subset1 <- subset |> def_cols(index = .I) |> arrange_cols(at="start", index)
subset1
```

```{r}
# arrange_cols
new_dt <- arrange_cols(dt, at=`vehicle model`, qsec, vs, am)
new_dt
new_dt <- arrange_cols(dt, at="start", c("qsec", "vs", "am"))
new_dt
new_dt <- arrange_cols(dt, at="mpg", qsec:am, carb)
new_dt
new_dt <- arrange_cols(dt, at="start", qsec:am, mpg)
new_dt
```

