---
title: "Wrapper functions for data.table"
author: "kleu046"
date: "2021-08-22"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{function_details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>"
)
```

```{r setup}
library(wr.data.table)
```

# Introduction

While ***data.table*** offers fast functions to tidy and summarize data, the syntax may often can often be hard to remember if not used or practised frequently.

This package aims to provide *wrapper functions* to carry out common tasks using functions in the data.table library without having to remember the data.table syntax.

This vignette outlines detail usages for these wrapper function

## data.table.withRownames()

This function converts a data.frame into a data.table and use the row names as the first column.

This is equivalent to:

> data.table(df, keep.rownames = True)

For examples, the *mtcars* contains car manufacturers and models as row names.

```{r data.table.withRownames, echo=TRUE}
dt_withRownames <- data.table.withRownames(mtcars)

head(dt_withRownames)
```

if a *data.table* is provided as arugment to this function, it does nothing but returning the input.

Inputs other than data.frame or data.table will result in an error.
